Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-03-17 VegOzone.tex
Second draft of results

The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing temperature acclimation $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplified effect of light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species. Drought effects (SWP) do only matter if a bespoke GS is not taken into account.

In more detail for downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Downy birch displays the largest range of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light threshold is especially pronounced, e.g. $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2019. Without bespoke GS 2018 stands out as a higher ozone damage risk year compared to 2019. However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type. SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased. Norway spruce displays a medium sensitivity to the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarly large bias as downy birch. We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type. Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

The estimated biomass reduction in \unit{\%} for the bespoke temperature response function of type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}. The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations in light threshold and SWP=on. For downy birch and perennial grassland, substantial reductions in total biomass are predicted for any of the two years. Norway spruce is least affected. The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland. For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated. In comparison with average results for the default MM parameterization (with bespoke GS), we find a non negligible (above systematic uncertainties) low-bias for both downy birch and perennial grassland. For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range.

@@

---
The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
\begin{itemize}
\item Increasing temperature acclimation $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplified effect of light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species.
 The impact of the timing of the GS can be at the same order of magnitude as the acclimation of response function but points into the opposite direction. Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak induces large uncertainties in anticipation of an acclimation to colder climates.

>>>>

The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFT and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species. Drought effects (SWP) do only matter if a bespoke GS is not taken into account. We will look at the specific species or PFTs in more detail in the following.

For downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Downy birch displays the largest variance of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light threshold is especially pronounced, e.g. $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the boreal type without bespoke GS in 2019. Without bespoke GS, 2018 stands out as the year with the higher ozone damage risk than 2019. However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type. SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased. Norway spruce displays a medium sensitivity to the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarly large bias as downy birch. We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type. Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

The estimated biomass reduction in \unit{\%} for the bespoke temperature response function of type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}. The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations of the light threshold and SWP=on. For downy birch and perennial grassland, substantial reductions in total biomass are predicted for any of the two years. Norway spruce is least affected. The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland. For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated. In comparison with average results for the default MM parameterization (with bespoke GS), we find a non negligible (above systematic uncertainties) low-bias for both downy birch and perennial grassland. For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range, this may reflect the more comprehensive research on this species or PFT.

@@

---
The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFTs and between years. Thus predicted ozone uptake is XSlarger than for MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species.
 The impact of the timing of the GS can be at the same order of magnitude as the acclimation of response function but points into the opposite direction. Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak induces large uncertainties in anticipation of an acclimation to colder climates.
